---
title: "climwin_demo"
author: "LRA"
date: "2/4/2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE)
library(tidyverse)
library(lme4)
library(climwin)
library(lubridate)
library(sjPlot)
library(performance)
```

## Demo for Coffee & Code

Use the climwin pacakage to identify climate windows of interest for further modeling
Requires 2 data sets - response variable and climate data
package is structured to use daily climate data
Response variable here is seagrass shoot density

# Prep biological data 
```{r biodata}
# read in environmental data at transect level
env19 <- read_csv("data/all_survey_metrics_transect_2019.csv")
env20 <- read_csv("data/all_survey_metrics_transect_2020.csv")
env21 <- read_csv("data/all_survey_metrics_transect_2021.csv")

env19 <- select(env19,-c("PrevalenceMean","SeverityMean","LesionAreaMean"))
env20 <- select(env20,-c("PrevalenceMean","SeverityMean","LesionAreaMean", "EpiphytePerAreaSe"))
env21 <- select(env21,-c("PrevalenceMean","SeverityMean","LesionAreaMean", "EpiphytePerAreaSe"))
env_c <- rbind(env19,env20, env21)

# read in disease data at blade level
disease19 <- read_csv("data/disease_2019.csv")
disease20 <- read_csv("data/disease_metrics_2020.csv")
disease21 <- read_csv("data/disease_metrics_2021.csv")

disease19 <- select(disease19, c("SampleDate","SampleId","Region","SiteCode","TidalHeight", "Transect", "Blade",
                                 "Year", "Prevalence", "LesionArea", "BladeArea", "Severity"))
disease20 <- select(disease20, c("SampleDate","SampleId","Region","SiteCode","TidalHeight", "Transect", "Blade",
                                 "Year", "Prevalence", "LesionArea", "BladeArea", "Severity"))
disease21 <- select(disease21, c("SampleDate","SampleId","Region","SiteCode","TidalHeight", "Transect", "Blade",
                                 "Year", "Prevalence", "LesionArea", "BladeArea", "Severity"))

dis_c <- rbind(disease19,disease20, disease21)

dis_summ <- dis_c %>%
  group_by(Year,Region, SiteCode,Transect, SampleDate) %>%
  summarise(TransectPrevalence=mean(Prevalence), TransectLesionArea=mean(LesionArea), TransectSeverity=mean(Severity),
            TransectBladeArea=mean(BladeArea))

dis_env <- full_join(dis_summ, env_c, by=c("Year","Region","SiteCode","Transect"))
# limit to observations with no missing data
dis_env <- select(dis_env, c("SampleDate","Region","SiteCode","TidalHeight","Transect","TransectPrevalence",
                              "TransectLesionArea","TransectBladeArea","TransectSeverity","Year",
                             "DensityShootsMean", "CountBlades"))
dis_env <- na.omit(dis_env)
# add IDs
dis_env$Meadow <- paste(dis_env$Region, dis_env$SiteCode, sep = "_")
head(dis_env)
```

## Prep climate data

```{r clim data, echo=FALSE}
# use daily data?
mur <- read_csv("data/MUR_daily.csv")
meadows <- tibble(Meadow=unique(mur$Meadow))
# convert dates for climwin
mur$Day <- factor(format(mur$Date, "%d/%m/%Y"))
head(mur)
```

## Fit baseline model

```{r baseline}
# limit disease data to MUR sites
dis_env <- left_join(meadows,dis_env, by="Meadow")
# scale continuous variables based on restricted data set
dis_env$sBladeArea <- scale(dis_env$TransectBladeArea, center=TRUE,scale=TRUE)
dis_env$sDensityShootsMean <- scale(dis_env$DensityShootsMean, center=TRUE, scale=TRUE)
# convert dates for climwin
dis_env$SampleDay <- factor(format(dis_env$SampleDate, "%d/%m/%Y"))
# convert year to factor
dis_env$fYear <- as.factor(dis_env$Year)

fit_den <- lmer(DensityShootsMean ~ sBladeArea + TidalHeight + fYear + TransectPrevalence +
                  (1|Region) + (1|Meadow),
                data = dis_env)
summary(fit_den)
```

## Look for climate windows

```{r climwin}

clim_mur_den_tr <- slidingwin(xvar = list(TempAnomaly = mur$DiffMean, 
                                           TempAnomalyHeat = mur$DiffMeanHeat, 
                                           Temp = mur$analysed_sst),
                               cdate = mur$Day,
                               bdate = dis_env$SampleDay,
                               baseline = fit_den,
                               cohort = as.factor(dis_env$Year),
                               cinterval = "month",
                               range = c(6, 0),
                               type = c("absolute","relative"),
                               refday = c(20, 7),
                               cmissing = "method1",
                               stat = c("mean"),
                               func = "lin",
                               spatial = list(dis_env$Meadow, mur$Meadow))
clim_mur_den_tr$combos
# best models include terms for TempAnomalyHeat for month opening 4 months before July 20 and closing 3 months before and 
# Temp Anomaly Heat for 1 month window opening 5 months prior and closing 4 months prior to sampling
# summary(clim_mur_den_tr[[5]]$BestModel)
plotdelta(dataset = clim_mur_den_tr[[5]]$Dataset)
plot_model(clim_mur_den_tr[[5]]$BestModel, show.values = TRUE, show.p = TRUE)
```

## Check for overfitting with randomization

```{r random}
# check with randomization
# clim_mur_den_tr_randomized <- randwin(repeats=100,
#                                       xvar = list(TempAnomaly = mur$DiffMean, 
#                                                   TempAnomalyHeat = mur$DiffMeanHeat, 
#                                                   Temp = mur$analysed_sst),
#                                       cdate = mur$Day,
#                                       bdate = dis_env$SampleDay,
#                                       baseline = fit_den,
#                                       cohort = as.factor(dis_env$Year),
#                                       cinterval = "month",
#                                       range = c(6, 0),
#                                       type = c("absolute","relative"),
#                                       refday = c(20, 7),
#                                       cmissing = "method1",
#                                       stat = c("mean"),
#                                       func = "lin",
#                                       spatial = list(dis_env$Meadow, mur$Meadow))
clim_mur_den_tr_randomized <- readRDS("output/climwin_density_daily_MUR_transect.rds")
clim_mur_den_tr_randomized[[5]]
plothist(dataset=clim_mur_den_tr[[5]]$Dataset, datasetrand = clim_mur_den_tr_randomized[[5]])
pvalue(dataset=clim_mur_den_tr[[5]]$Dataset, datasetrand = clim_mur_den_tr_randomized[[5]], metric="AIC")
```

## Update baseline model to include climate

```{r new_mod}
head(clim_mur_den_tr[[5]]$BestModelData)
# climate predictor is for best model is Temp Anomaly Heat for 1 month window opening 5 months prior and closing 4 months prior to sampling
# rather than recalculate, use this dataset but rename yvar and climate
new_data <- clim_mur_den_tr[[5]]$BestModelData
new_data <- rename(new_data, DensityShootsMean=yvar, TempAnomalyHeat=climate)
fit_den_update <- lmer(DensityShootsMean ~ sBladeArea + TidalHeight + fYear + TransectPrevalence+ TempAnomalyHeat +
                  (1|Region) + (1|Meadow),
                data = new_data)
summary(fit_den_update)
AIC(fit_den, fit_den_update)
```

